### 2021-07-02 Analysis/02_Simulation_pipeline.Rmd
Update 02_Simulation_pipeline.Rmd

---
title: "Simulation pipeline"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

This script is the repeatable script for measuring changes in the different diversity metrics and how they are affected by the different stressors.

## Requirements for the simulations

In order to run the simulations (on your machine, cluster, etc...) you'll need to load the following in your current environment:

```{r, eval = FALSE}
## Getting the non-CRAN (yet) libraries
library(devtools)
devtools::install_github("TGuillerme/dads")
```

```{r}
## Libraries
require(dads)
require(dispRity)
require(BAT)
require(FD)
require(ks)
require(TPD)
```

> Note that the libraries do not need to be loaded (i.e. no need for `library()`) but needs to exist.

You'll also need the following custom functions (located in [here on github](https://github.com/TGuillerme/eco_metrics_simulations/tree/master/Functions))

```{r}
## Loading the functions
source("../Functions/melodic.rao.R")
source("../Functions/make.simulation.pipeline.R")
source("../Functions/simulation.pipeline.R")
```

## Setting up the simulations

You then need to set up the simulation parameters using `make.simulation.pipeline()`.
You can check the manual of that function by opening the function to see the different options but we suggest you use the default to replicate the parameters used in the manuscript.

By default, the simulation does:

 * Create a tree from a pure birth process with 2 independent birth-death processes to make a 2D trait space. The speciation (birth) rate is set to 1 and the simulation stops when reaching 200 taxa.
 * Randomly remove 20%, 40%, 60% and 80% of the data to simulate the random (null) stressor.
 * Remove 20%, 40%, 60% and 80% of the data using the required `type` of stressor (see below and/or `01_Simulating_spaces_and_stressors.Rmd`).
 * Measure five families of metrics on the randomly reduced and stressor reduced spaces: dendrogram, dissimilarity, convex hull, kernel hypervolume and probability density based metrics.
 * All the results are then collected into a table with the two stressors as rows and the different 15 metrics as columns.

You then just need to set up which type of stressor you are going to simulate the metrics with:

```{r}
## Creating the function for simulating competition
simulate.competition <- make.simulation.pipeline(type = "competition")
```

You can then just run one replicate of the simulation by simply calling that function like (not run here):

```{r, eval = FALSE}
## Running one replicate
simulate.competition()
```

## Running the simulations

In order to simulate enough replicates we will replicate the simulation pipeline $n$ times until adding 5% more replicates does not increase the overall variance by more than 5% for any metric.
In other words, we replicated each simulation $n$ times where $n$ is estimated so that $var(n + 0.05 \times n) < 0.05 \times var(n)$.

For that we will use the custom `optim.replicate()` function which comes with one diagnosis function (checking if the variance for each metric is < 0.05) and one summarising (summarising the output of `simulate.competition()` for diagnosis) function

```{r}
## Loading the optimisation functions
source("../Functions/optim.replicate.R")
source("../Functions/optim.helpers.R")
source("../Functions/analyse.replicates.R")
```

You can then run this replicate function that will start by running 20 replicates (`mimimum`) and will increase the number of replicates by 5% until reaching 1000 replicates (`maximum`) or until the addition 5% increase does not increase the variance for each metric with each different level of removal by less than 5% (120 parameters!).

There are options to

```{r, eval = TRUE}
## Running the full optimisation
results <- optim.replicate(input.fun = simulate.competition,
                           diagnose = var,
                           summarise = optim.summarise,
                           minimum = 20, maximum = 1000,
                           stop.variance = 0.05,
                           verbose = TRUE,
                           bkp.path = "../Data/Processed/",
                           bkp.name = "sim_competition.rda")
```
> Note that you can use the parallel version by adding the argument `parallel = TRUE` that is automatised using the package `future` and `future.apply` which makes the whole process run on all your available cores but is way less verbose. Without using the parallel version, it takes roughly 10 hours (overnight) to run the analyses on an old 2.2Ghz 8GB core.

## Checking the results

We can then visualise the "convergence" of the results using the `analyse.replicate` function:

```{r, eval = FALSE}
source("../Functions/analyse.replicates.R")
## Check the change in variance for all 120 parameters
analyse.replicates("../Data/Processed/sim_filtering.rda") #140
analyse.replicates("../Data/Processed/sim_equalizing.rda") #155
analyse.replicates("../Data/Processed/sim_facilitation.rda") #126
analyse.replicates("../Data/Processed/sim_competition.rda")
```