### 2024-11-28 Manuscript/Tables/results_differences2d_1.tex
Added all results tables

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & EQ.d\_1 & EQ.ses\_1 & EQ.d\_2 & EQ.ses\_2 & EQ.d\_3 & EQ.ses\_3 & EQ.d\_4 & EQ.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 66.11*** & 4.1 & 68.86*** & 3.87 & 60.36*** & 3 & 43.83*** & 2.01 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 43.52*** & 4.02 & 41.95*** & 3.83 & 34.67*** & 2.89 & 23.62*** & 1.85 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 36.44*** & 4.88 & 41.74*** & 5.17 & 39.04*** & 4.21 & 30.16*** & 2.99 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 31.3*** & 5.32 & 37.36*** & 4.98 & 33.84*** & 3.53 & 23.2*** & 2.13 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.11*** & 3.48 & 0.05*** & 1.64 & 0.01*** & 0.44 & -0.01*** & -0.25 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 2.06*** & 4.98 & 1.78*** & 4.39 & 1.38*** & 3.23 & 0.92*** & 2.06 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.24*** & 2.01 & 0.17*** & 1.51 & 0.11*** & 1.03 & 0.06*** & 0.59 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 2.81*** & 20.32 & 2.07*** & 18.96 & 1.43*** & 15.61 & 0.82*** & 11.23 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0.09*** & 4.35 & 0.04*** & 2.12 & 0.01*** & 0.7 & -0.01*** & -0.49 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.33*** & -5.46 & -0.24*** & -4.94 & -0.16*** & -3.81 & -0.09*** & -2.12 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0 & -0.01 & 0* & -0.18 & 0*** & -0.27 & 0*** & -0.23 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & -0.02*** & -0.69 & -0.03*** & -1.32 & -0.03*** & -1.58 & -0.03*** & -1.54 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (2D). EQ = equalising; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}