### 2024-11-28 Manuscript/Tables/results_differences2d_3.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & FI.d\_1 & FI.ses\_1 & FI.d\_2 & FI.ses\_2 & FI.d\_3 & FI.ses\_3 & FI.d\_4 & FI.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 44.79*** & 2.45 & 42.48*** & 1.98 & 36.94*** & 1.58 & 25.79*** & 1.08 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 31.28*** & 2.26 & 27.45*** & 1.85 & 22.78*** & 1.45 & 14.93*** & 0.95 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 28.59*** & 3.14 & 28.96*** & 2.67 & 26.06*** & 2.22 & 18.98*** & 1.58 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 21.29*** & 3.08 & 24.42*** & 2.89 & 22.75*** & 2.08 & 15.6*** & 1.39 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.04*** & 0.95 & 0.01 & 0.14 & -0.01** & -0.26 & -0.02*** & -0.41 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 1.27*** & 2.4 & 1.03*** & 1.88 & 0.81*** & 1.48 & 0.53*** & 0.98 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.14*** & 1.34 & 0.09*** & 0.92 & 0.06*** & 0.62 & 0.03*** & 0.33 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 1.2*** & 4.88 & 0.99*** & 5.02 & 0.78*** & 4.56 & 0.49*** & 4.4 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0.06*** & 2.38 & 0.04*** & 1.71 & 0.03*** & 1.1 & 0.01*** & 0.46 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.05*** & -0.71 & -0.02*** & -0.4 & -0.01*** & -0.23 & 0 & -0.07 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0* & 0.25 & 0. & 0.17 & 0 & 0.11 & 0 & -0.03 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.02*** & 0.46 & 0.01*** & 0.34 & 0 & -0.03 & 0*** & -0.2 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (2D). FI = filtering; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}