### 2024-11-28 Manuscript/Tables/results_differences4d_2.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & FA.d\_1 & FA.ses\_1 & FA.d\_2 & FA.ses\_2 & FA.d\_3 & FA.ses\_3 & FA.d\_4 & FA.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 517.78*** & 0.82 & 783.51*** & 0.96 & 877.91*** & 0.9 & 724.23*** & 0.67 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 940.19*** & 0.71 & 1134.97*** & 0.72 & 1169.29*** & 0.68 & 945.03*** & 0.53 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 235.02*** & 1.11 & 401.13*** & 1.49 & 477.06*** & 1.45 & 439.74*** & 1.23 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 29.27*** & 0.74 & 51.57*** & 1.85 & 57.78*** & 2.4 & 43.32*** & 2.14 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0 & -0.03 & 0 & 0.05 & 0.01*** & 0.21 & 0*** & 0.1 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.36*** & 0.47 & 0.4*** & 0.53 & 0.39*** & 0.52 & 0.31*** & 0.43 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.03*** & 0.4 & 0.03*** & 0.37 & 0.03*** & 0.33 & 0.02*** & 0.24 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 0.53*** & 1.38 & 0.62*** & 2.38 & 0.6*** & 3.1 & 0.46*** & 3.63 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0.03*** & 2.88 & 0.02*** & 2.69 & 0.02*** & 2.31 & 0.01*** & 0.93 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.03*** & -0.75 & -0.03*** & -1 & -0.03*** & -1.14 & -0.02*** & -0.92 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0. & -0.21 & 0*** & -0.75 & 0*** & -0.52 & 0*** & -0.4 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.22*** & 4.28 & 0.11*** & 3.5 & 0.03*** & 1.38 & -0.01*** & -0.55 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (4D). FA = facilitation; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}