### 2024-11-28 Manuscript/Tables/results_differences4d_4.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & CO.d\_1 & CO.ses\_1 & CO.d\_2 & CO.ses\_2 & CO.d\_3 & CO.ses\_3 & CO.d\_4 & CO.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 160.54*** & 0.28 & 239.24*** & 0.28 & 239.86*** & 0.23 & 181.09*** & 0.15 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 436.95*** & 0.3 & 491.69*** & 0.28 & 445.45*** & 0.24 & 307.31*** & 0.15 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 115.67*** & 0.61 & 153.13*** & 0.48 & 159.12*** & 0.42 & 123.17*** & 0.27 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 6.44*** & 0.69 & 8.26*** & 0.61 & 8.72*** & 0.52 & 6.84*** & 0.36 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.04*** & 0.93 & 0.03*** & 0.91 & 0.02*** & 0.7 & 0.01*** & 0.44 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.28*** & 0.41 & 0.24*** & 0.34 & 0.19*** & 0.28 & 0.13*** & 0.18 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.03*** & 0.35 & 0.02*** & 0.26 & 0.02*** & 0.23 & 0.01*** & 0.14 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 0.68*** & 2.23 & 0.59*** & 2.72 & 0.44*** & 2.8 & 0.28*** & 2.85 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0*** & 0.39 & 0.01*** & 0.61 & 0.01*** & 0.61 & 0*** & 0.39 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.01*** & -0.43 & -0.01*** & -0.38 & -0.01*** & -0.3 & 0*** & -0.22 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0 & -0.08 & 0 & -0.02 & 0 & -0.02 & 0 & -0.02 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.01*** & 0.48 & 0.01*** & 0.5 & 0.01*** & 0.48 & 0*** & 0.21 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (4D).CO = competition; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}