### 2024-11-28 Manuscript/Tables/results_differences8d_1.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & EQ.d\_1 & EQ.ses\_1 & EQ.d\_2 & EQ.ses\_2 & EQ.d\_3 & EQ.ses\_3 & EQ.d\_4 & EQ.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 761.11*** & 1.34 & 962.64*** & 1.11 & 873.68*** & 0.81 & 544.78*** & 0.43 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 1031482.51*** & 0.39 & 1415980.4*** & 0.33 & 1316657.33*** & 0.23 & 798612.09*** & 0.12 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 7348.88*** & 2.06 & 29098.27*** & 1.93 & 48535.88*** & 1.7 & 47957.63*** & 1.11 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 52.07*** & 3.25 & 62.28*** & 3 & 55*** & 2.41 & 33.6*** & 1.35 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.08*** & 2.07 & 0.05*** & 1.37 & 0.03*** & 0.72 & 0.01*** & 0.29 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 2*** & 2.46 & 1.49*** & 1.83 & 1.07*** & 1.27 & 0.62*** & 0.73 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.12*** & 1.6 & 0.08*** & 1.16 & 0.06*** & 0.85 & 0.03*** & 0.5 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 6.09*** & 12.85 & 4.25*** & 12.1 & 2.83*** & 10.04 & 1.56*** & 7.82 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0*** & 0.47 & 0*** & 0.43 & 0*** & 0.22 & 0 & -0.01 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.07*** & -4.87 & -0.03*** & -3.81 & -0.02*** & -2.76 & -0.01*** & -1.71 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0*** & -0.72 & 0*** & -0.42 & 0*** & -0.29 & 0*** & -0.25 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.02*** & 0.86 & 0.01*** & 0.47 & 0*** & 0.25 & 0 & 0.05 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D). EQ = equalising; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}