### 2024-11-28 Manuscript/Tables/results_differences8d_2.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & FA.d\_1 & FA.ses\_1 & FA.d\_2 & FA.ses\_2 & FA.d\_3 & FA.ses\_3 & FA.d\_4 & FA.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 514.85*** & 1.18 & 555.45*** & 0.78 & 600.74*** & 0.78 & 443.13*** & 0.49 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 877212.48*** & 0.53 & 1040106.12*** & 0.36 & 1253555.6*** & 0.35 & 961196.13*** & 0.22 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 5448.5*** & 0.89 & 18489.32*** & 0.91 & 35772.76*** & 1.48 & 39487.59*** & 1.06 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 63.57*** & 1.5 & 83.62*** & 1.59 & 97.82*** & 4.02 & 75.03*** & 2.69 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & -0.02*** & -0.54 & -0.01*** & -0.4 & -0.01*** & -0.22 & 0*** & -0.11 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.13*** & 0.17 & 0.14*** & 0.17 & 0.14*** & 0.18 & 0.13*** & 0.16 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.02*** & 0.39 & 0.01*** & 0.24 & 0.01*** & 0.22 & 0.01*** & 0.16 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 0.49*** & 0.97 & 0.44*** & 1.4 & 0.43*** & 1.91 & 0.38*** & 2.35 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0.02*** & 3.95 & 0.02*** & 2.45 & 0.01*** & 1.65 & 0.01*** & 0.72 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & 0* & -0.25 & -0.01*** & -0.72 & 0*** & -0.77 & 0*** & -0.69 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0*** & -1.09 & 0*** & -1.26 & 0*** & -1.19 & 0*** & -0.67 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.25*** & 7.86 & 0.13*** & 4.81 & 0.05*** & 2.64 & 0.01*** & 0.64 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D). FA = facilitation; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}