### 2024-11-28 Manuscript/Tables/results_differences8d_3.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & FI.d\_1 & FI.ses\_1 & FI.d\_2 & FI.ses\_2 & FI.d\_3 & FI.ses\_3 & FI.d\_4 & FI.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 512.45*** & 0.98 & 601.87*** & 0.74 & 538.56*** & 0.54 & 373.2*** & 0.32 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 759454.01*** & 0.51 & 1036742.44*** & 0.41 & 984207.67*** & 0.3 & 724993.83*** & 0.18 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 6282.8*** & 1.77 & 22928.75*** & 1.72 & 33761.59*** & 1.38 & 35235.64*** & 0.95 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 54.91*** & 3.53 & 71.72*** & 3.02 & 67.56*** & 2.39 & 46.86*** & 1.62 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.02*** & 0.45 & 0.01* & 0.15 & 0.01* & 0.15 & 0** & 0.12 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.66*** & 0.77 & 0.45*** & 0.56 & 0.33*** & 0.43 & 0.22*** & 0.29 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.06*** & 0.8 & 0.03*** & 0.5 & 0.02*** & 0.32 & 0.01*** & 0.19 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 1.73*** & 1.99 & 1.19*** & 1.89 & 0.86*** & 2.12 & 0.57*** & 2.4 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0.02*** & 1.8 & 0.01*** & 1.45 & 0.01*** & 1.29 & 0*** & 0.52 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.01*** & -0.79 & 0*** & -0.57 & 0*** & -0.49 & 0*** & -0.32 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0*** & -0.96 & 0*** & -0.95 & 0*** & -0.73 & 0*** & -0.5 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.13*** & 1.73 & 0.07*** & 1.59 & 0.03*** & 1.41 & 0.01*** & 0.52 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D). FI = filtering; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}