### 2024-11-28 Manuscript/Tables/results_differences8d_4.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & CO.d\_1 & CO.ses\_1 & CO.d\_2 & CO.ses\_2 & CO.d\_3 & CO.ses\_3 & CO.d\_4 & CO.ses\_4 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 79.64*** & 0.14 & 130.26*** & 0.16 & 105.45*** & 0.1 & 92.81*** & 0.08 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 67526.58*** & 0.04 & 173367.22*** & 0.06 & 216655.44*** & 0.06 & 150636.07*** & 0.04 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 1312.02*** & 0.3 & 6855.69*** & 0.47 & 8664.07*** & 0.32 & 8093.46*** & 0.2 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 3.79*** & 0.29 & 7.28*** & 0.44 & 6.82*** & 0.32 & 5.38*** & 0.23 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.02*** & 0.54 & 0.02*** & 0.5 & 0.01*** & 0.4 & 0.01*** & 0.3 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.21*** & 0.27 & 0.21*** & 0.28 & 0.16*** & 0.21 & 0.1*** & 0.13 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.01*** & 0.18 & 0.01*** & 0.2 & 0.01*** & 0.14 & 0.01*** & 0.11 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 0.6*** & 1.4 & 0.61*** & 2.07 & 0.47*** & 2.15 & 0.32*** & 2.17 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0 & 0.11 & 0** & 0.23 & 0*** & 0.37 & 0*** & 0.27 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & 0** & -0.29 & 0*** & -0.54 & 0*** & -0.34 & 0*** & -0.21 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0 & -0.03 & 0 & -0.1 & 0 & 0.05 & 0 & -0.1 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.01* & 0.24 & 0.01*** & 0.31 & 0*** & 0.25 & 0 & 0.07 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D).CO = competition; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}