### 2024-11-28 Manuscript/Tables/results_differencesempirical.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:22:09 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllll}
  \hline
 & empirical.d\_1 & empirical.ses\_1 & empirical.d\_2 & empirical.ses\_2 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & -7.28*** & -1.53 & -4.8*** & -1.08 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & -44.19*** & -1.74 & -9.72*** & -0.37 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & -17.07*** & -2.29 & -3.07*** & -0.55 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & -6.04*** & -2.38 & -4.32*** & -1.78 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.02*** & 1.45 & -0.03*** & -1.63 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.11*** & 1.28 & -0.05*** & -0.44 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.03*** & 1.64 & -0.03*** & -1.19 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & -0.41*** & -1.91 & -0.16*** & -0.52 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0* & 0.24 & 0.01*** & 2.35 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & 0.01*** & 0.97 & 0.02*** & 1.06 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0*** & -0.72 & 0*** & 0.69 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & -0.03*** & -2.23 & -0.02*** & -0.7 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each
stressor (empirical); d = difference, s = standardised effect size. Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1.} 
\end{table}