### 2024-11-28 Manuscript/Tables/results_model2d.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:16:05 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & equalizing.Slope & equalizing.adj.R\verb|^|2 & facilitation.Slope & facilitation.adj.R\verb|^|2 & filtering.Slope & filtering.adj.R\verb|^|2 & competition.Slope & competition.adj.R\verb|^|2 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 0.017*** & 0.02 & -0.015** & 0.012 & 0.026*** & 0.057 & -0.001 & -0.001 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 0.014** & 0.011 & -0.016** & 0.014 & 0.024*** & 0.038 & -0.009* & 0.007 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 0.007 & 0.002 & -0.006 & 0 & 0.03*** & 0.069 & 0.004 & 0 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 0.002 & -0.001 & -0.055*** & 0.07 & 0.001 & -0.001 & -0.02*** & 0.048 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.185*** & 0.578 & 0.022** & 0.009 & 0.102*** & 0.117 & 0.022** & 0.013 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.08*** & 0.396 & 0.009. & 0.003 & 0.054*** & 0.195 & 0.008* & 0.005 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.018** & 0.009 & -0.023*** & 0.04 & 0.014* & 0.004 & -0.022*** & 0.048 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 0.066*** & 0.436 & -0.028*** & 0.046 & -0.009. & 0.004 & -0.022*** & 0.038 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0.227*** & 0.572 & 0.044*** & 0.037 & 0.091*** & 0.155 & 0.002 & -0.001 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.09*** & 0.293 & -0.002 & -0.001 & -0.027*** & 0.043 & -0.011* & 0.008 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & 0.018*** & 0.049 & 0.046*** & 0.179 & 0.004 & 0 & -0.012*** & 0.02 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.09*** & 0.356 & 0.14*** & 0.28 & 0.022*** & 0.037 & -0.009* & 0.005 \\ 
   \hline
\end{tabular}
\caption{Results of the model scaled metric ~ removal level per stressor (2D)} 
\end{table}