### 2024-11-28 Manuscript/Tables/results_model4d.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Nov 28 15:16:05 2024
\begin{table}[ht]
\scriptsize
\centering
\begin{tabular}{rllllllll}
  \hline
 & equalizing.Slope & equalizing.adj.R\verb|^|2 & facilitation.Slope & facilitation.adj.R\verb|^|2 & filtering.Slope & filtering.adj.R\verb|^|2 & competition.Slope & competition.adj.R\verb|^|2 \\ 
  \hline
Functional evenness\\distances\\\texttt{FD::dbFD()\$FEve} & 0.054*** & 0.143 & -0.009 & 0.002 & 0.032*** & 0.076 & 0.002 & 0 \\ 
  Functional evenness\\dendrogram\\\texttt{BAT::evenness} & 0.013* & 0.007 & -0.032*** & 0.092 & 0.015** & 0.012 & -0.002 & 0 \\ 
  Functional evenness\\hypervolume\\\texttt{BAT::kernel.evenness} & 0.049*** & 0.122 & 0.008 & 0.002 & 0.061*** & 0.194 & 0.018*** & 0.038 \\ 
  Functional evenness\\probability density\\\texttt{TPD::REND} & 0.03*** & 0.135 & -0.01 & 0.002 & 0.024*** & 0.09 & 0.001 & -0.001 \\ 
  Rao's quadratic entropy\\distances\\\texttt{FD::dbFD()\$RaoQ} & 0.164*** & 0.539 & -0.02** & 0.012 & 0.039*** & 0.023 & -0.002 & -0.001 \\ 
  Functional dispersion\\dendrogram\\\texttt{BAT::dispersion} & 0.108*** & 0.535 & -0.009** & 0.009 & 0.067*** & 0.258 & 0.012*** & 0.025 \\ 
  Functional dispersion\\hypervolume\\\texttt{BAT::kernel.dispersion} & 0.064*** & 0.145 & -0.004 & 0.002 & 0.045*** & 0.05 & 0.005* & 0.004 \\ 
  Divergence\\probability density\\\texttt{TPD::REND} & 0.11*** & 0.52 & -0.03*** & 0.071 & 0.034*** & 0.065 & 0.004 & 0.001 \\ 
  Alpha diversity\\dendrogram\\\texttt{BAT::alpha} & 0.153*** & 0.393 & 0.059*** & 0.083 & 0.014* & 0.005 & -0.035*** & 0.048 \\ 
  Functional richness\\convex hull\\\texttt{BAT::hull.alpha} & -0.118*** & 0.517 & 0.013*** & 0.018 & -0.025*** & 0.06 & -0.002 & 0 \\ 
  Functional richness\\hypervolume\\\texttt{BAT::kernel.alpha} & -0.008** & 0.009 & -0.015*** & 0.016 & -0.006 & 0.002 & -0.001 & -0.001 \\ 
  Functional richness\\probability density\\\texttt{TPD::REND} & 0.016*** & 0.037 & 0.196*** & 0.535 & 0.014*** & 0.02 & -0.001 & -0.001 \\ 
   \hline
\end{tabular}
\caption{Results of the model scaled metric ~ removal level per stressor (4D)} 
\end{table}