### 2025-01-23 Manuscript/Tables/results_differences2d_2.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:55:47 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & FA.d\_rm1 & FA.ses\_rm1 & FA.d\_rm2 & FA.ses\_rm2 & FA.d\_rm3 & FA.ses\_rm3 & FA.d\_rm4 & FA.ses\_rm4 \\ 
  \hline
Functional richness\\ probability density\\ \texttt{TPD::REND} & 34.1*** & 1.51 & 32.08*** & 1.19 & 32.28*** & 1.14 & 28.07*** & 0.98 \\ 
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 21.27*** & 1.29 & 18.72*** & 1.03 & 17.59*** & 0.96 & 14.25*** & 0.77 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 21.81*** & 1.94 & 21.56*** & 1.56 & 22.13*** & 1.53 & 19.58*** & 1.33 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 22.76*** & 2.2 & 26.13*** & 1.63 & 29.21*** & 2.18 & 24.35*** & 1.85 \\ 
  Divergence\\ probability density\\ \texttt{TPD::REND} & 0 & -0.03 & 0 & 0.1 & -0.01** & -0.16 & -0.01*** & -0.2 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 0.76*** & 1.12 & 0.6*** & 0.89 & 0.55*** & 0.82 & 0.46*** & 0.69 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.08*** & 0.77 & 0.06*** & 0.59 & 0.05*** & 0.47 & 0.03*** & 0.33 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 0.88*** & 3.28 & 0.7*** & 3.57 & 0.6*** & 4.16 & 0.44*** & 4.18 \\ 
  Functional evenness\\ probability density\\ \texttt{TPD::REND} & 0.02*** & 0.61 & 0.01*** & 0.53 & 0.01*** & 0.31 & 0 & -0.09 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & -0.09*** & -1.01 & -0.06*** & -0.84 & -0.05*** & -0.81 & -0.04*** & -0.73 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0** & 0.35 & 0 & 0.09 & 0*** & -0.24 & 0*** & -0.45 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.25*** & 5.31 & 0.15*** & 3.43 & 0.06*** & 2.12 & -0.01*** & -0.45 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (2D). FA = FA; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}