### 2025-01-23 Manuscript/Tables/results_differences2d_4.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:55:52 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & CO.d\_rm1 & CO.ses\_rm1 & CO.d\_rm2 & CO.ses\_rm2 & CO.d\_rm3 & CO.ses\_rm3 & CO.d\_rm4 & CO.ses\_rm4 \\ 
  \hline
Functional richness\\ probability density\\ \texttt{TPD::REND} & 17.17*** & 0.79 & 18.21*** & 0.76 & 16.88*** & 0.68 & 12.58*** & 0.48 \\ 
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 13.24*** & 0.89 & 14.64*** & 0.96 & 12.41*** & 0.81 & 9.06*** & 0.57 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 11.24*** & 1.1 & 12.26*** & 1.05 & 11.73*** & 0.95 & 8.88*** & 0.67 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 7.71*** & 1.17 & 11*** & 1.24 & 11.32*** & 1.14 & 8.68*** & 0.74 \\ 
  Divergence\\ probability density\\ \texttt{TPD::REND} & 0.04*** & 0.87 & 0.04*** & 1.01 & 0.03*** & 0.72 & 0.02*** & 0.49 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 0.51*** & 0.9 & 0.51*** & 0.93 & 0.41*** & 0.77 & 0.31*** & 0.56 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.06*** & 0.58 & 0.05*** & 0.52 & 0.04*** & 0.4 & 0.03*** & 0.26 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 0.71*** & 3.18 & 0.67*** & 4.05 & 0.53*** & 4.31 & 0.36*** & 3.85 \\ 
  Functional evenness\\ probability density\\ \texttt{TPD::REND} & 0.04*** & 1.3 & 0.04*** & 1.73 & 0.03*** & 1.25 & 0.02*** & 0.83 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & -0.02** & -0.29 & -0.01. & -0.13 & 0 & -0.08 & 0 & -0.01 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0** & 0.32 & 0*** & 0.35 & 0*** & 0.3 & 0*** & 0.19 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.03*** & 0.76 & 0.03*** & 0.99 & 0.01*** & 0.62 & 0.01*** & 0.32 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (2D).CO = CO; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}