### 2025-01-23 Manuscript/Tables/results_differences4d_1.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:55:55 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & EQ.d\_rm1 & EQ.ses\_rm1 & EQ.d\_rm2 & EQ.ses\_rm2 & EQ.d\_rm3 & EQ.ses\_rm3 & EQ.d\_rm4 & EQ.ses\_rm4 \\ 
  \hline
Functional richness\\ probability density\\ \texttt{TPD::REND} & 1024.68*** & 2.29 & 1305.5*** & 1.94 & 1221.81*** & 1.37 & 794.58*** & 0.74 \\ 
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 2277.38*** & 2.33 & 2433.1*** & 2.04 & 2101.05*** & 1.48 & 1349.76*** & 0.84 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 449.97*** & 3.43 & 720.41*** & 3.62 & 780.96*** & 2.77 & 628.64*** & 1.8 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 42.72*** & 4.14 & 51.54*** & 3.7 & 46.87*** & 2.64 & 30.12*** & 1.5 \\ 
  Divergence\\ probability density\\ \texttt{TPD::REND} & 0.13*** & 3.66 & 0.07*** & 1.91 & 0.03*** & 0.92 & 0.01*** & 0.16 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 2*** & 3.53 & 1.65*** & 2.84 & 1.25*** & 2.06 & 0.78*** & 1.23 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.16*** & 1.76 & 0.11*** & 1.25 & 0.08*** & 0.88 & 0.04*** & 0.52 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 4.19*** & 15.4 & 2.92*** & 15.04 & 1.97*** & 11.6 & 1.1*** & 8.48 \\ 
  Functional evenness\\ probability density\\ \texttt{TPD::REND} & 0.01*** & 1.2 & 0*** & 0.37 & 0*** & -0.21 & 0*** & -0.55 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & -0.23*** & -5.95 & -0.14*** & -5.16 & -0.08*** & -4.15 & -0.04*** & -2.33 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0*** & -0.56 & 0*** & -0.4 & 0*** & -0.49 & 0** & -0.12 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & -0.01** & -0.33 & -0.01*** & -0.68 & -0.01*** & -0.62 & -0.01*** & -0.54 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (4D). EQ = equalising; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}