### 2025-01-23 Manuscript/Tables/results_differences4d_3.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:56:00 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & FI.d\_rm1 & FI.ses\_rm1 & FI.d\_rm2 & FI.ses\_rm2 & FI.d\_rm3 & FI.ses\_rm3 & FI.d\_rm4 & FI.ses\_rm4 \\ 
  \hline
Functional richness\\ probability density\\ \texttt{TPD::REND} & 667.99*** & 1.31 & 871.01*** & 1.14 & 808.97*** & 0.83 & 549.73*** & 0.48 \\ 
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 1625.13*** & 1.34 & 1706.46*** & 1.14 & 1401.89*** & 0.83 & 908.74*** & 0.49 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 398.17*** & 2.59 & 587.5*** & 2.44 & 586.06*** & 1.89 & 437.87*** & 1.12 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 33.97*** & 3.24 & 40.35*** & 3.04 & 36.06*** & 1.83 & 23.54*** & 1.03 \\ 
  Divergence\\ probability density\\ \texttt{TPD::REND} & 0.05*** & 1.29 & 0.04*** & 1.08 & 0.02*** & 0.67 & 0.01*** & 0.31 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 1.16*** & 1.84 & 0.91*** & 1.4 & 0.67*** & 1.01 & 0.42*** & 0.62 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.1*** & 1.15 & 0.07*** & 0.83 & 0.04*** & 0.55 & 0.02*** & 0.32 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 1.83*** & 4.43 & 1.42*** & 4.4 & 1*** & 3.98 & 0.59*** & 3.23 \\ 
  Functional evenness\\ probability density\\ \texttt{TPD::REND} & 0.01*** & 1 & 0.01*** & 1.18 & 0.01*** & 0.85 & 0*** & 0.34 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & -0.04*** & -1.03 & -0.02*** & -0.87 & -0.01*** & -0.49 & -0.01*** & -0.25 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0* & -0.26 & 0 & -0.14 & 0 & 0.1 & 0 & 0.07 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.02*** & 0.88 & 0.01*** & 0.54 & 0* & 0.18 & 0 & 0.09 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (4D). FI = FI; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}