### 2025-01-23 Manuscript/Tables/results_differences8d_1.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:56:13 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & EQ.d\_rm1 & EQ.ses\_rm1 & EQ.d\_rm2 & EQ.ses\_rm2 & EQ.d\_rm3 & EQ.ses\_rm3 & EQ.d\_rm4 & EQ.ses\_rm4 \\ 
  \hline
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 1031482.51*** & 0.39 & 1415980.4*** & 0.33 & 1316657.33*** & 0.23 & 798612.09*** & 0.12 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 7348.88*** & 2.06 & 29098.27*** & 1.93 & 48535.88*** & 1.7 & 47957.63*** & 1.11 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 52.07*** & 3.25 & 62.28*** & 3 & 55*** & 2.41 & 33.6*** & 1.35 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 2*** & 2.46 & 1.49*** & 1.83 & 1.07*** & 1.27 & 0.62*** & 0.73 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.12*** & 1.6 & 0.08*** & 1.16 & 0.06*** & 0.85 & 0.03*** & 0.5 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 6.09*** & 12.85 & 4.25*** & 12.1 & 2.83*** & 10.04 & 1.56*** & 7.82 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & -0.07*** & -4.87 & -0.03*** & -3.81 & -0.02*** & -2.76 & -0.01*** & -1.71 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0*** & -0.72 & 0*** & -0.42 & 0*** & -0.29 & 0*** & -0.25 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.02*** & 0.86 & 0.01*** & 0.47 & 0*** & 0.25 & 0 & 0.05 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D). EQ = equalising; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}