### 2025-01-23 Manuscript/Tables/results_differences8d_2.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:56:18 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & FA.d\_rm1 & FA.ses\_rm1 & FA.d\_rm2 & FA.ses\_rm2 & FA.d\_rm3 & FA.ses\_rm3 & FA.d\_rm4 & FA.ses\_rm4 \\ 
  \hline
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 877212.48*** & 0.53 & 1040106.12*** & 0.36 & 1253555.6*** & 0.35 & 961196.13*** & 0.22 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 5448.5*** & 0.89 & 18489.32*** & 0.91 & 35772.76*** & 1.48 & 39487.59*** & 1.06 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 63.57*** & 1.5 & 83.62*** & 1.59 & 97.82*** & 4.02 & 75.03*** & 2.69 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 0.13*** & 0.17 & 0.14*** & 0.17 & 0.14*** & 0.18 & 0.13*** & 0.16 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.02*** & 0.39 & 0.01*** & 0.24 & 0.01*** & 0.22 & 0.01*** & 0.16 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 0.49*** & 0.97 & 0.44*** & 1.4 & 0.43*** & 1.91 & 0.38*** & 2.35 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & 0* & -0.25 & -0.01*** & -0.72 & 0*** & -0.77 & 0*** & -0.69 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0*** & -1.09 & 0*** & -1.26 & 0*** & -1.19 & 0*** & -0.67 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.25*** & 7.86 & 0.13*** & 4.81 & 0.05*** & 2.64 & 0.01*** & 0.64 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D). FA = FA; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}