### 2025-01-23 Manuscript/Tables/results_differences8d_3.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:56:25 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & FI.d\_rm1 & FI.ses\_rm1 & FI.d\_rm2 & FI.ses\_rm2 & FI.d\_rm3 & FI.ses\_rm3 & FI.d\_rm4 & FI.ses\_rm4 \\ 
  \hline
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 759454.01*** & 0.51 & 1036742.44*** & 0.41 & 984207.67*** & 0.3 & 724993.83*** & 0.18 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 6282.8*** & 1.77 & 22928.75*** & 1.72 & 33761.59*** & 1.38 & 35235.64*** & 0.95 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 54.91*** & 3.53 & 71.72*** & 3.02 & 67.56*** & 2.39 & 46.86*** & 1.62 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 0.66*** & 0.77 & 0.45*** & 0.56 & 0.33*** & 0.43 & 0.22*** & 0.29 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.06*** & 0.8 & 0.03*** & 0.5 & 0.02*** & 0.32 & 0.01*** & 0.19 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 1.73*** & 1.99 & 1.19*** & 1.89 & 0.86*** & 2.12 & 0.57*** & 2.4 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & -0.01*** & -0.79 & 0*** & -0.57 & 0*** & -0.49 & 0*** & -0.32 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0*** & -0.96 & 0*** & -0.95 & 0*** & -0.73 & 0*** & -0.5 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.13*** & 1.73 & 0.07*** & 1.59 & 0.03*** & 1.41 & 0.01*** & 0.52 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D). FI = FI; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}