### 2025-01-23 Manuscript/Tables/results_differences8d_4.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:56:30 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & CO.d\_rm1 & CO.ses\_rm1 & CO.d\_rm2 & CO.ses\_rm2 & CO.d\_rm3 & CO.ses\_rm3 & CO.d\_rm4 & CO.ses\_rm4 \\ 
  \hline
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 67526.58*** & 0.04 & 173367.22*** & 0.06 & 216655.44*** & 0.06 & 150636.07*** & 0.04 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 1312.02*** & 0.3 & 6855.69*** & 0.47 & 8664.07*** & 0.32 & 8093.46*** & 0.2 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 3.79*** & 0.29 & 7.28*** & 0.44 & 6.82*** & 0.32 & 5.38*** & 0.23 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 0.21*** & 0.27 & 0.21*** & 0.28 & 0.16*** & 0.21 & 0.1*** & 0.13 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.01*** & 0.18 & 0.01*** & 0.2 & 0.01*** & 0.14 & 0.01*** & 0.11 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 0.6*** & 1.4 & 0.61*** & 2.07 & 0.47*** & 2.15 & 0.32*** & 2.17 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & 0** & -0.29 & 0*** & -0.54 & 0*** & -0.34 & 0*** & -0.21 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0 & -0.03 & 0 & -0.1 & 0 & 0.05 & 0 & -0.1 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.01* & 0.24 & 0.01*** & 0.31 & 0*** & 0.25 & 0 & 0.07 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (8D).CO = CO; d = difference, s = standardised effect size; 1 to 4 correspond to the levels of data removal (20\%, 40\%, 60\% and 80\%). Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}