### 2025-01-23 Manuscript/Tables/results_differencesempirical.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:56:35 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllll}
  \hline
 & empirical.d\_rm1 & empirical.ses\_rm1 & empirical.d\_rm2 & empirical.ses\_rm2 \\ 
  \hline
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & -44.19*** & -1.74 & -9.72*** & -0.37 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & -17.07*** & -2.29 & -3.07*** & -0.55 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & -6.04*** & -2.38 & -4.32*** & -1.78 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 0.11*** & 1.28 & -0.05*** & -0.44 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.03*** & 1.64 & -0.03*** & -1.19 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & -0.41*** & -1.91 & -0.16*** & -0.52 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & 0.01*** & 0.97 & 0.02*** & 1.06 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0*** & -0.72 & 0*** & 0.69 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & -0.03*** & -2.23 & -0.02*** & -0.7 \\ 
   \hline
\end{tabular}
\caption{Average difference between the null metric and the stressed metric (raw) for each level of removal and each stressor (empirical); d = difference, s = standardised effect size. Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1. Standardised effect size was calculated using the Hedges' \textit{g} with a correction for small-sample bias.} 
\end{table}