### 2025-01-23 Manuscript/Tables/results_model2d.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Jan 23 18:48:58 2025
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{rllllllll}
  \hline
 & equalizing.Slope & equalizing.adj.R\verb|^|2 & facilitation.Slope & facilitation.adj.R\verb|^|2 & filtering.Slope & filtering.adj.R\verb|^|2 & competition.Slope & competition.adj.R\verb|^|2 \\ 
  \hline
Functional richness\\ probability density\\ \texttt{TPD::REND} & 0.02*** & 0.02 & -0.02** & 0.01 & 0.03*** & 0.06 & 0 & 0 \\ 
  Functional richness\\ hypervolume\\ \texttt{BAT::kernel.alpha} & 0.01* & 0 & -0.02*** & 0.02 & 0.02*** & 0.03 & -0.01*** & 0.01 \\ 
  Functional richness\\ convex hull\\ \texttt{BAT::hull.alpha} & 0.01 & 0 & -0.01 & 0 & 0.03*** & 0.06 & 0 & 0 \\ 
  Alpha diversity\\ dendrogram\\ \texttt{BAT::alpha} & 0 & 0 & -0.06*** & 0.07 & 0 & 0 & -0.02*** & 0.05 \\ 
  Divergence\\ probability density\\ \texttt{TPD::REND} & 0.19*** & 0.58 & 0.02** & 0.01 & 0.1*** & 0.12 & 0.02** & 0.01 \\ 
  Functional dispersion\\ hypervolume\\ \texttt{BAT::kernel.dispersion} & 0.08*** & 0.39 & 0.01. & 0 & 0.05*** & 0.19 & 0.01. & 0 \\ 
  Functional dispersion\\ dendrogram\\ \texttt{BAT::dispersion} & 0.02* & 0.01 & -0.02*** & 0.04 & 0.01. & 0 & -0.02*** & 0.05 \\ 
  Rao's quadratic entropy\\ distances\\ \texttt{FD::dbFD()\$RaoQ} & 0.06*** & 0.39 & -0.03*** & 0.05 & -0.01* & 0.01 & -0.02*** & 0.04 \\ 
  Functional evenness\\ probability density\\ \texttt{TPD::REND} & 0.23*** & 0.57 & 0.04*** & 0.04 & 0.09*** & 0.16 & 0 & 0 \\ 
  Functional evenness\\ hypervolume\\ \texttt{BAT::kernel.evenness} & -0.09*** & 0.29 & 0 & 0 & -0.03*** & 0.04 & -0.01* & 0.01 \\ 
  Functional evenness\\ dendrogram\\ \texttt{BAT::evenness} & 0.02*** & 0.05 & 0.05*** & 0.18 & 0 & 0 & -0.01*** & 0.02 \\ 
  Functional evenness\\ distances\\ \texttt{FD::dbFD()\$FEve} & 0.09*** & 0.36 & 0.14*** & 0.28 & 0.02*** & 0.04 & -0.01* & 0 \\ 
   \hline
\end{tabular}
\caption{Results of the model scaled metric $\sim$ removal level per stressor (2D)} 
\end{table}