### 2025-02-02 Manuscript/ecography_reponse_letter_2.tex
response updates

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Response to reviewers},
  pdfauthor={Thomas Guillerme},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{Response to reviewers}
\author{Thomas Guillerme}

\begin{document}
\maketitle

Dear Editor,

We are grateful for the detailed and useful comments from all two reviewers
We have integrated changes associated with the suggestion into the manuscript and responded to them in detail below
For clarity we have displayed the \textcolor{blue}{reviewers' comments in blue} and our own responses in black
We fully describe each change below.
We have also attached a tracked change version of the manuscript to hopefully improve clarifying our changes.

Best regards,

Thomas Guillerme, on behalf of all co-authors.

% \textcolor{blue}{Recommendation by the Subject Editor (Dr. Barry Brook):}

% \textcolor{blue}{While Reviewer 1 is largely satisfied with the improvements and only requests minor revisions (e.g., clarifying a few sentences, adding references, and fixing small typographical issues), Reviewer 2 identifies more substantive points that need further attention—particularly around your approach to comparing metrics across different dimensionalities, the conceptual definition of single vs. multiple traits, and the impact of your second scaling step on the observed patterns. Clarifying why and how you simulate trait dimensions and potentially standardising them at four dimensions for TPD comparisons, seem to be important for interpretability.}

% \textcolor{blue}{In your revision,

% - please focus on resolving these methodological and conceptual concerns to improve transparency and strengthen the presentation of your core findings.

% - You will also need to address the referees’ points regarding:
%   - the consistency of dimension usage,
%   - the rationale for simulating “multidimensional” traits,
%   - and the potential biases introduced by your second scaling step.

% Additionally, please follow Reviewer 1’s minor editorial suggestions to refine readability. }

\textcolor{blue}{\textbf{Reviewer: 1}}

\textcolor{blue}{Thank you for dedicating your time and attention to addressing the comments. Your manuscript has significantly improved in readability, content, discussion, and interpretations. The addition of the text box at the end is particularly appreciated, as it effectively highlights the key message of the manuscript. At this stage, I only recommend minor revisions, and after those are addressed, I recommend the manuscript for publication. Below I provided my minor comments and edits to the manuscript.}

\textcolor{blue}{Line 89-90: I think that some references are necessary here.}

We changed this sentence and merged it with the next one after discovering some previously missed literature:

\textit{Although some previous work has been done in understanding patterns in the context of mechanisms and processes (e.g. Novack-Gottshall (2016a,b))), we argue that often the pattern description follows some previously used framework without a proper evaluation of its adequacy to answer the how and why.} Lines 89-92. %@@@

\textcolor{blue}{Line 175-177: This sentence is not very clear. I recommend rephrasing.}

We rephrased the sentence to:

\textit{In other words, pairs of observations that share similar trait combinations are more common in the trait space than observations with dissimilar trait combinations (i.e. observations adjacent to each other in the trait space are more common).} Lines 175-178. %@@@

\textcolor{blue}{Line 183: Remove “-“.}

We replaced the dash by a comma.

\textcolor{blue}{Line 196: I don’t think you need the brackets when citing the function. }

We removed the brackets.

\textcolor{blue}{Line 201: Please explain why you chose p = 3.}

We've now explained why we chose p=3:

\textit{here we arbitrarily used $p=3$ which was a relatively low scaling value that still resulted in visible changes observable in 2D} Lines 202-203.% @@@.

\textcolor{blue}{Line 322: Correct “de Hawaiian species”.}

We fixed the type.

\textcolor{blue}{Line 339: I would recommend placing figure 2 after the first results paragraph.}
\textcolor{blue}{Figure 3: I would recommend placing figure 3 in the results section. }

The placing of figures is done automatically here by the compiler.
But we'll make sure to forward this suggestion to the copy editors after the manuscript is accepted.

\textcolor{blue}{Line 388: Remove “say”.}

We removed the word ``say''.

\textcolor{blue}{Line 426-432: Please rephrase and simplify.}

We've now simplified the sentence to:

\textit{Our results show that the loss of  functional diversity for several metrics was not more than expected by chance.
This could be due to the taxonomic distribution of the extinct birds: these Hawaiian data comprise approximatively $58$\% passerines while the majority of anthropogenic extinctions have involved non-passerines (approximatively $75$\%).
Non-passerines tend to be larger and possess more unique morphological traits (thus resulting in relatively large reductions in functional diversity).} Lines 431-436. %@@@

\textcolor{blue}{Line 542: Please specify what suggestions you are referring to.}

We've now specified the following:

\textit{For example, we can follow the suggestions from Mammola et al. (2021) highlighting which kind of methodological pipeline best supports which type of metric of interest (Table 3).} Lines 545-547. %@@@

\textcolor{blue}{\textbf{Reviewer: 3}}

\textcolor{blue}{I greatly appreciate the effort the authors have put into revising the manuscript.
The substantial reformulation has significantly improved the understanding of processes, mechanisms, and expectations.
I particularly value the expanded explanation of how the stressors are applied.
However, I have a few concerns that I believe should be addressed to further enhance the clarity and methodological rigor of the manuscript.
First, the comparison of metrics across different dimensionalities raises questions, particularly when metrics from lower-dimensional spaces (e.g., the TPD package) are compared to those measured in higher-dimensional spaces.
Given the manuscript's aim to compare metrics measured in a standardized way, ensuring consistency across dimensions is fundamental.
Additionally, this version of the manuscript has highlighted concerns regarding the simulation of traits—specifically, whether the approach involves a single trait with multiple dimensions or multiple independent traits.
This distinction has important implications for the assumptions of independence and variance, and subsequently, for the simulated patterns within the trait space.
Below, I provide specific comments and suggestions to address these points.
I will recommend a further revision to clarify the selection and implementation of these approaches, as this will enhance the transparency and methodological rigor of the manuscript.}

\textcolor{blue}{Specific comments: }

\textcolor{blue}{Lines 128-135: The capacity of the algorithm to simulate the stressor strongly depends on the distribution of the data and the assumption of independence in the simulated dimensions (as should be the case in the simulation example).
Given this, it is unclear why a single trait is simulated as multidimensional.
This approach feels unnecessarily complex, potentially confounding the reader and introducing covariation that should be tested.
To clarify, a trait is typically defined as a single measurable characteristic of an organism, particularly when considering the definition referenced in Violle et al. (2007).
The concept of multidimensionality in trait spaces usually applies when multiple traits are combined into a matrix, with each column representing a distinct trait or a principal component summarizing trait variation.
However, it is not common practice to describe a single trait using multiple variables (i.e., columns or dimensions).}

\textcolor{blue}{Regarding the statement in Lines 132–135:
\textit{“These aspects are often expressed in one dimension but can in fact be described in any number of dimensions. For example, for trait defined as “leaf insertion angle”, this can measured in one dimension (an angle in degrees) or three dimensions (the same angle expressed as the trigonometric relation of three sides of a triangle).”}
While this explanation clarifies the authors’ definition of a “multidimensional trait,” it appears to complicate the methodology.
If traits are often expressed in a single dimension (e.g., angle in degrees), I do not see the necessity to simulate them as “a 2, 4 or 8 dimensional trait, in other words a Brownian Motion simulated in a 2, 4 or 8 dimensions.” (lines 128-129 If I understood the text correctly, this approach results in a trait space built around a single trait defined by multiple dimensions.}

\textcolor{blue}{Instead, it would be more intuitive and broadly applicable to simulate 2, 4, or 8 separate traits.
These traits, together, would define a 2-, 4-, or 8-dimensional trait space.
Since these traits would be uncorrelated, they would correspond directly to the same number of principal components, aligning with the definition of dimensions provided in Supplementary Table 1.
Moreover, simulating a single multidimensional trait may inadvertently introduce covariation across dimensions.
For example, using “leaf insertion angle” as an illustration, the trigonometric relationships of the triangle’s sides and angles inherently interconnect mathematically (e.g., through sine, cosine, and tangent).
This interconnection could introduce correlation in the variance, even under Brownian assumptions.
While this correlation may not be inherently incorrect or distinct from real ecological data, the assumption of independence in the simulation should be explicitly tested.
Correlated dimensions might create clustering in the trait space, potentially biasing the application of stressors in the simulation context. 
Finally, simulating independent traits using Brownian motion without fixing the variance could also address the issue of homogeneity of variance across dimensions (Lines 463–464).
Homogeneity of variance may introduce similar bias in the simulated mechanisms and it is far from approximating a set of independent trait or dimensions.
I belive that simulating independent trait may mitigate that concern.}

We have changed the description of the trait to multiple uni-dimensional traits.
We used (and maybe insisted on) the previous definition of a single uncorrelated multidimensional trait because the way they can be defined in the \texttt{treats} package (a trait = any number of dimensions as long as the process is the same).
We changed the description to the following:

\textit{First, we simulated multiple independent random time dependent traits (Brownian Motion) under a model where lineages only speciate (no extinction; i.e. a pure birth speciation model) until reaching 200 observations in \texttt{R} (R Core Team, 2024) using \texttt{treats} (Guillerme, 2024).
We simulated either 2, 4 or 8 independent (uncorrelated) Brownian Motion traits (equivalent to 2, 4 or 8 dimensional traits).} Lines 124-124. %@@@

However, we kept the broader definition of trait.
Violle et al. (2007) provided an ``unambiguous definition of plant trait, with a particular emphasis on functional trait'', which as per the authors aims was proposed to clarify and and conceptualise the definition on functional traits (``a surrogate of organismal performance'') specifically in plant ecology.
Here we build on McGill et al. (2006)'s more generalised approach, and, more interestingly on the work of Dawson et al. (2021) of how ecologists \textit{actually} operationalise the concept of trait based on researchers understanding vs. proposed definitions.
We propose through the paragraph on lines 124-139 to expand that to ``any measurable aspect of an organism'' which we hope allows expansion beyond one dimensional traits and, more importantly, beyond ecology.
I.e. we believe that this definition of trait more accurately also includes how traits are used in microbiology (e.g. bacterial colony shape), macroevolution (e.g. 3D geometric morphometrics) or in palaeontology (e.g. ``position of a specific foramen in a skull relative to another feature'').

\textcolor{blue}{Lines 191 – 209: This paragraph includes parts that require clarification.
It begins by stating that the 'algorithm reduces the probability of resampling observations in regions of the trait space that have many observations and increases it in regions that have few observations', which corresponds to the equation ix(1-b)p.
However, if this algorithm is intended to approximate a competition mechanism ‘where observations in dense regions of the trait space are more likely to be removed than in sparse regions of the trait space.’ and ‘We chose this algorithm to approximate a stressor that could increase the probability of extinction for observations that share traits combination’ shouldn’t the pattern be the reverse?
Furthermore, the formula, by subtracting 1 from the probability, seems to skew sampling towards less populated regions of the space, which, if I understand correctly, does not align with the concept of competitive exclusion.}

We believe the confusion stems from what probability we are talking here.
In a way here we approximate both competitive exclusion (i.e. increased ``extinction'' in densely populated regions) the same way as increased ``survival'' in sparsely populated regions.
The reviewer is true that these two mechanisms are fundamentally different in ecology but we believe that for our approximations, they result in the same patterns: extinction in dense region (due to high competition) and survival in sparse regions (due to low competition).

The formula illustrates how we measure the probability of being sampled which is decreasing in densely populated regions.
This probability of being sampled reflects the probability of ``surviving'', i.e. of not being removed from the simulation.
We've now added this precision throughout the explanation:

\textit{(or, inversely, observations are more likely to be ``preserved''/``surviving'' in less dense regions)}. Lines 208-209. %@@@

\textcolor{blue}{Lines 280-289: It seems that I may need further clarification regarding the second step of the scaling.
If the data are only divided by the maximum value, the results would generally have a maximum of 1 in cases of consistently negative or positive metrics across all levels of species exclusion (e.g., -2/-2 = 1 and +2/+2 = 1).
Similarly, when some metrics have negative values and others are positive (across different levels of removal), the range would not necessarily be constrained between -1 and 1 unless all values fall within the range [0, 1] (e.g., functional richness might not range between -1 and 1).}

We did not use any metrics that output negative values.
This means that the scaled and null centered values are always ranging between -1 and 1 in theory (even some values might not be achieved in practice).
Hopefully the raw results in the supplementary materials now highlight this.

\textcolor{blue}{Furthermore, I am concerned that this second scaling could reduce or bias the patterns of the metrics, potentially affecting the conclusions of the manuscript.
For instance, in Supplementary Figure 8 (simulated data for 4 dimensions), taking FD:FEve for facilitation as a reference, subtracting the MetricStressor from the MetricNull results in a relatively linear pattern, with a general decrease in evenness as species removal increases (if I am seeing well from the figure), which aligns with expectations under this mechanism.
However, in Figure 2, the same metric exhibits a more drastic decrease in stressor evenness at 60\% and an increase at 80\%, challenging to reconcile with the expected mechanism, given that the data were simulated to approximate a facilitation mechanism.
A similar trend can be observed across other examples (e.g., Supplementary Figures with raw data), which may explain the contrasting patterns between historical and extant empirical data, with some metrics showing fewer extreme differences than observed.
I suggest further exploring the impact of this second scaling on the final results.}

We are not entirely sure how to handle this comment.
If we understood correctly, the reviewer suggests we use a different null scaling method to see how that changes the results for specific metrics.
Doing so will certainly result in different results.
Although we believe that there is no need to test a range of scaling methods since the null centering one was chosen specifically to make the results more readable in this specific paper: does metric A captures a ``directional'' change in a specific mechanism induced pattern and is this change distinguishable from a null mechanism induced pattern?

We are definitely sure that there are more possibilities of metric combinations, pattern simulations, results summarising that would be of interest to readers but with our limited time resources, we decide to chose only the specific one presented in this manuscript.
Thankfully though, we made this whole project entirely reproducible so hopefully some worker wishing to pursue our work could easily add and modify what we did here.

\textcolor{blue}{Figure 1: To enhance readability, I suggest including a legend for the colours (orange and blue) directly in the figure, not just in the caption.
I appreciate the improvements made to expand the figure.
However, if metrics C to F are scaled for the null mechanisms, I assume that in the first two panels of section G, the term 'scaled' could be removed, as they likely represent the actual metric used to compute the scaled metric in panel 3.
If the metrics are further scaled, this should be clarified in the text.
For completeness and readability, I also suggest explicitly adding the details of the second scaling performed relative to the four levels of removal.}

We've now added the colour legend to the plot figure and removed the mention to the scaled metrics in the legend.
As requested by this reviewer, we now only display the raw metrics scored (in the supplementaries) and the scaled and null centered metrics scores (in the manuscript and in the supplementaries).

\textcolor{blue}{Lines 330 – 335/ Figure 4: I appreciate the clarification regarding the TPD metric.
However, I regret to say that comparing a metric built on four dimensions with one working on five dimensions does not seem appropriate, particularly given the aims of the manuscript.
Considering the amount of variation explained by the first four dimensions (Supplementary Figure 11) and the manuscript's focus on exploring metric performance with a real dataset rather than depicting bird diversity after extinction, standardizing the dimensionality across all metrics is necessary.
This approach would enable a fair comparison, even if it requires a slight compromise on the 95\% variation threshold, targeting approximately 90\% instead (as observed in the histogram for four dimensions, which aligns with the minimum number of dimensions recommended by Pigot et al. 2020).
This adjustment is particularly important, as metric performance may vary significantly across different dimensionalities (e.g., metrics evaluated across 2, 4, or 8 dimensions).
Standardizing to four dimensions would also ensure consistency with the dimensionality used in the simulated dataset and maintain uniformity throughout the analyses.
Regarding the 8-dimensional space, since it is not central to the analysis, I suggest removing the TPD approach for this case to avoid introducing unnecessary variability, while clearly justifying this decision as a limit of the package in the text.}

The aim of including the empirical data was not to compare the metrics directly \textit{per se} but rather to provide a second example that readers preferring real world data can hopefully relate to.
In line with the aim, we have thus decided to keep the empirical dataset to 5 dimensions, namely to hopefully link it to a paper that is currently in preparation by some of the authors (with different objectives but a same subset of the data).
Furthermore, reducing the dataset to 4 dimensions would require rerunning the full analyses which will require CPU time that we currently don't have available because of other projects using our allocated computational time.

We do agree with the reviewers point that including the \texttt{TPD} metrics that are limited to 4 dimensions can be misleading.
We have now removed all the curves for the \texttt{TPD} metrics for any space of more than four dimensions (including the empirical space).
We added this footnote to Table 3:

\textit{-The \texttt{TPD} package currently only allows calculations for traitspaces up to four dimensions. This metric was exclude from traitspaces with more than four dimensions.}

\textcolor{blue}{Furthermore, upon focusing on Supplementary Figure 11, I have a concern.
Based on the description in the main text, I understand that the analyses are performed starting from a single trait space (comprising 118 points), where species are removed either randomly or based on actual extinction events, and metrics are then recalculated.
This approach seems appropriate for evaluating how species are distributed and how metric performance changes under different scenarios.
However, the figure comparing extant and pre-1500 species leads me to wonder whether the trait spaces were recalculated based on the subset of species following extinctions.
Am I correct in assuming that the analyses are based on a single trait space constructed with all 118 points?}

Yes, the empirical trait space is based on 118 observations split between three groups: 37 species present on the island today, 63 species that were present on island in 1500 and 18 species that went extinct before 1500.
We believe that the confusion here might come from the equivalence between species present (63 at 1500 and 37 in the present) and species absent (55 (46\%) at 1500 and 81 (68\%) in the present).
We have tried now to clarify this in the caption of the figure 3 and in the one from the supplementary figure concerning the dimensionality selection.

\textcolor{blue}{Discussion: While I appreciate the revisions and the inclusion of the Caveats section, I feel that Box 2 might not add substantial value, as it primarily refers to previously published materials rather than the manuscript's main findings.
Expanding the Results/Discussion section instead could be more beneficial.
For instance, the simulations were performed across 2, 4, and 8 dimensions, and the metrics and their relationships seem to vary across these spaces with different dimensionalities.
Addressing this aspect in greater detail would add significant value to the manuscript and better align with its stated aims.}

We believe that this specific point conflicts with reviewer 1 (``The addition of the text box at the end is particularly appreciated, as it effectively highlights the key message of the manuscript'' - reviewer 1 above).
In this instance we thus prefer keeping the box 2 since we believe it might add value to at least some readers (and hopefully doesn't distract from the reading for other readers).
Furthermore, it is not entirely clear to us what the reviewer meant with the metrics relationship ``varying across spaces with different dimensionalities''.
Did they meant that spaces with different dimensionalities affect different metrics differently?
We did mention the effect of the curse in the methods section:

\textit{This curse changes the properties of space occupancy in a non-linear way depending on each observation's distribution and the number of dimensions.} lines 251-253.
We now also mention it in the discussion:

\textit{Note also that, as mentioned above, different dimensionalities can have different effects on the metrics, depending on both the traitspace and the metric nature (Bellman et al., 1957).} Lines 388-390. %@@@

\textcolor{blue}{Supplementary Table 1: This table is a valuable addition to the manuscript.
However, geographical location or landmark coordinates should not be considered a trait of an observation, particularly when applying the definition of a trait as used throughout the text (e.g., Violle et al. 2007 and others).
Additionally, I suggest revising the terminology in the same column to avoid defining a 'trait' as a 'trait'; using terms like 'characteristic' or 'aspect of observations' might align better with the main text's definitions.
Finally, I recommend considering the inclusion of this table in the main text, as it enhances the reader's understanding of the concepts presented.}

We've now added the the glossary table (supplementary table 1) to the main text as suggested with the definition of trait redefined as a characteristic (good catch!).
We did however keep our broader definition of a trait as opposed to Violle et al (2007)'s (see response to first comment of Reviewer 3 above).
We have added this precision in the table:

\textit{Note that our definition of a trait is a generalisation ofViolle et al. (2007); McGill et al.
(2006); Dawson et al. (2021)’s suggestions which is expanded to include traits beyond the ones used for functional ecology (macroevolution, palaeontology, microbiology, etc.).} At the end of Table 1.

\textcolor{blue}{Supplementary Figures 7–10: I appreciate the addition of the new figures showing the raw metrics.
However, given the range of the values (particularly for functional richness) and the legend indicating that the metrics are scaled, it is unclear to me what the values are referencing.
I suggest not scaling the metrics in this case, as they should naturally correspond to the same units as the null values, ensuring clarity and consistency.
This adjustment would enhance the interpretability of the data.}

We have now updated all these supplementary figures to show the raw results with the scale bars indicating the range of values for each metric.

\textcolor{blue}{Supplementary Tables 6–18: I suggest including the measured metrics in addition to the differences currently reported.
Additionally, I noticed that some mechanisms (e.g., Tables 8, 10, 12, 14, and 16) show differences that warrant double-checking.
For example, if the reported values are raw, the differences for functional divergence with TPD should be constrained between 0 and 1, as the metric itself is bounded within this range.
However, some values in the tables exceed this limit.}

Good spotting, the row names where mismatched (reversed) with the rows actual content.
We have now fixed that in the supplementaries.

\textcolor{blue}{Finally, could you clarify how the SES is calculated in these cases?}

We used the Hedge's \textit{g} method to measure the standardised effect size (from the \texttt{effectsize} package) corrected for small sample sizes:

\begin{equation}
g = \frac{
        median_{1} - median_{2}
      }{
    \sqrt{
      \frac{
           (n_{1}-1)\sigma_{1} + (n_{2}-1)\sigma_{2}
          }{
            n_{1}+n_{2}-2
          }
        }
      }
\end{equation}

\textcolor{blue}{Lastly, I recommend a thorough review of the text, as I noticed several instances of missing punctuation throughout.}

We have double checked the manuscript again.
Hopefully there are no more typos.

\end{document}